Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-11 paper/paper.tex
bug fix.
minor

Figure~\ref{fig:mot-example} shows an example program and the state of
the heap at  the program point $\pi$, i.e. just  before the evaluation
of $(\length\  \pz)$.  The heap is  represented by a graph  in which a
node either  represents atomic  values ($\NIL$,  integers etc.),  or a
\CONS\ cell containing $\CAR$ and $\CDR$  fields, or a chunk of memory
representing  an  unevaluated  value   also  called  a  {\em  closure}
(represented  by  hatched  boxes).   Edges   in  the  graph  are  {\em
  references} and emanate from variables  or fields.  The figure shows
the  lists \px\  and  \pz\ partially  evaluated.   The evaluation  was
triggered    by   the    test   (\NULLQ~(\CAR~(\CDR~\pz)))    in   the
\SIF\ expression.

@@

The specific contributions of this work are
  \begin{itemize}
  \item  We formulate  a  liveness  analysis for  a  lazy first  order
    functional language. We show how  to compute function summaries in
    a  context  independent  manner  to achieve  a  context  sensitive
    analysis. We  further describe the transformation  of the liveness
    equations to grammar rules in a  way that the language accepted by
    the resulting grammar describes live heap locations.\mycomment{the
      proof of its correctness}
  \item We prove  the undecidability of the grammar  that results from
    the  analysis.   To overcome  this,  we  present an  algorithm  to
    transform grammar  to finite  state automata  (DFA) such  that the
    language accepted by the  DFA over-approximates the liveness (and
    hence a safe approximation).
  \item We have implemented a garbage  collector that can use the DFA
    to  retain live  cells that  also include  closures. We  have done
    extensive  experiments with  this garbage  collector and  obtained
    empirical results  that show  the effectiveness  of liveness-based
    garbage collection.
  \end{itemize}

@@

\bibliography{fun_hra}{}
\bibliographystyle{abbrv}
%\end{document}

>>>>

Figure~\ref{fig:mot-example} shows an example program and the state of
the heap at  the program point $\pi$, i.e. just  before the evaluation
of $(\length\  \pz)$.  The heap is  represented by a graph  in which a
node either  represents atomic  values ($\NIL$,  integers etc.),  or a
\CONS\ cell containing $\CAR$ and $\CDR$  fields, or a chunk of memory
representing  an  unevaluated  value   also  called  a  {\em  closure}
(represented  by  hatched  boxes).   Edges   in  the  graph  are  {\em
  references} and emanate from variables  or fields.  The figure shows
the  lists \px\  and  \pz\ partially  evaluated.   The evaluation  was
triggered    by   the    test   (\NULLQ~(\CAR~\pz))    in   the
\SIF\ expression.

@@

While an ideal garbage collector is  impractical, in this work we show
that  static analysis  of  heap data  can help  practical garbage
collectors in reclaiming more  garbage.  The specific contributions of
this paper are
\begin{itemize}
\item  We formulate  a  liveness  analysis for  a  lazy first  order
  functional language. We show how  to compute function summaries in
  a  context  independent  manner  to achieve  a  context  sensitive
  analysis. We  further describe the transformation  of the liveness
  equations to grammar rules in a  way that the language accepted by
  the resulting grammar describes live heap locations.\mycomment{the
    proof of its correctness}
\item We prove  the undecidability of the grammar  that results from
  the  analysis.   To overcome  this,  we  present an  algorithm  to
  transform grammar  to finite  state automata  (DFA) such  that the
  language accepted by the  DFA over-approximates the liveness (and
  hence a safe approximation).
\item We have implemented a garbage  collector that can use the DFA
  to  retain live  cells that  also include  closures. We  have done
  extensive  experiments with  this garbage  collector and  obtained
  empirical results  that show  the effectiveness  of liveness-based
  garbage collection.
\end{itemize}

@@

\bibliography{fun_hra}{}
\bibliographystyle{abbrv}
\end{document}
